# Compare areas of plate grown plants across treatments, genotypes and zygosity

```{r read-chunks, echo=FALSE, warning=FALSE}
suppressMessages(library(knitr))
suppressMessages(library(BCS.Phenotyping))
suppressMessages(library(reshape2))
suppressMessages(library(BCS.PhenoData))
suppressMessages(library(xlsx))
suppressMessages(library(plyr))
suppressMessages(library(markdown))
suppressMessages(library(ggplot2))
opts_chunk$set(fig.width=5, fig.height=5, tidy=TRUE, fig.align='center', 
               fig.show='hold', dev=c('png', 'pdf'), error=TRUE)
options(width=68)
options(markdown.HTML.options="mathjax")

meta <- as(path, "Oni")
path <- outdir(path, "BCS.Phenotyping")
df <- read.csv(file.path(path, "data.csv"), header=TRUE, stringsAsFactors=FALSE, row.names=1)
expdf <- create_plate_test_df(subset(df, !removed))
show(meta)
```

## Hypothesis testing

```{r create_phenodata, echo=FALSE}
traits <- colnames(expdf)[c(grep('^AREA', colnames(expdf)),
                            grep('^RGR', colnames(expdf)))]
vars <- intersect(c('treatment', 'reference', 'zygosity'), names(expdf))
nmeta <-
  suppressMessages(BCS.PhenoData::make.metadata(traits=traits,
                                                treats=vars,
                                                formats=rep("NUM.2", length(traits)),
                                                short=traits,
                                                units=c(rep("mm2", sum(grepl('^AREA',
                                                  colnames(expdf)))),
                                                  rep("", sum(grepl('^RGR', colnames(expdf)))))))
data <- data.frame(EXPT=expid(meta), GENOTYPE=expdf$GENOTYPE, BLOCK=expdf$BLOCK, 
                   SUBBLOCK=NA, SUBSUBBLOCK=NA, ROW=expdf$ROW, RANGE=expdf$RANGE, 
                   expdf[, c(vars, traits)])
data$TRT <- apply(data[,c('GENOTYPE', vars)], 1, paste, collapse='_')
pheno <- BCS.PhenoData::make.phenodata(data, nmeta)
```

We fit a linear model for the estimated areas, \\(y\\), according to the model

\\(y=\beta\_0b + \sum\_{i=1}\^{n}\left(\beta\_{i}w\_{i}\right) + \epsilon\\)

where \\(w\_{1\ldots{}n}\\) encode an identification vector corresponding to the levels of a factor created from a concatenation of the genotype, zygosity and treatment and \\(b\\) is the experiment block (replicate). We perform ANOVA for comparisons between homozygous and azygous plants to assess the null-hypotheses
\\(H\_{0\_{j,k}}: \beta\_j = \beta\_k\\)
for \\(j\\) and \\(k\\) corresponding to samples of the same genotype and treatment but different zygosity.

```{r anova, echo=FALSE}
warning('outlier removal is not yet implemented')
treats <- levels(factor(dataset(pheno)$treatment))
germs <- levels(factor(dataset(pheno)$GENOTYPE))
zygs <- levels(factor(dataset(pheno)$zygosity))
left <- data.frame(genotype=rep(germs, length(treats)),
                   treatment=rep(treats, each=length(germs)),
                   zygosity=zygs[2], stringsAsFactors=FALSE)
right <- data.frame(genotype=rep(germs, length(treats)),
                    treatment=rep(treats, each=length(germs)),
                    zygosity=zygs[1], stringsAsFactors=FALSE)

mat <-
  plyr::laply(1:nrow(left), function(i) c(paste(left[i,], collapse='_'),
                                          paste(right[i,], collapse='_')))
cm <- BCS.PhenoData::create.cm(contrasts=mat, levels=levels(factor(dataset(pheno)$TRT)))
results <- BCS.PhenoData::anova.phenodata(pheno, factor="TRT",  random="~ BLOCK", cm=cm)
results_df <- BCS.Phenotyping:::anova_phenodata_to_data_frame(results, pheno, left)

write.xlsx(results_df, 'anova-results.xlsx')
write.csv(results_df, 'anova-results.csv')
```

The test results are given as [anova-results.xlsx](anova-results.xlsx) and  [anova-results.csv](anova-results.csv)

We create effect plots for all comparsions where the mean is indicated by the dot, and the 95% confidence interval indicated by a horizontal line.

```{r effectplots, echo=FALSE, warning=FALSE, fig.with=7, fig.height=7}
for(i in 1:length(traits)){
  trait <- traits[i]
  res_df <- results[[trait]]
  colnames(res_df) <- gsub(paste(".", trait, sep = ""), "", colnames(res_df))
  res_df$Comp <- rownames(res_df)
  m <- max(c(abs(res_df$LCL), abs(res_df$UCL)), na.rm=TRUE) * 1.1
  txt <- paste("BCS.Base::effectplot.data.frame(res_df, unitname=\"Comp\", center=\"Effect\",
                                  lower=\"LCL\", upper=\"UCL\", labels.loc=\"above\",
                                  refline=0, col.ref=\"red\", 
                                  main=\"", trait,"\",
                                  xlim=c(", paste(-m,m, sep=','), "))")
  eval(parse(text=txt, n=-1))
}
```

## Session information
