# Sowing report

```{r read-chunks, echo=FALSE, warning=FALSE}
suppressMessages(library(rosettR))
opts_chunk$set(fig.width=5, fig.height=5, tidy=TRUE, fig.align='center',
               fig.show='hold', dev=c('png', 'pdf'), error=TRUE)
options(width=68)
options(markdown.HTML.options="mathjax")

meta <- readMeta('../..')
manifest <- readManifest('../..')
```

Plate phenotyping experiment with

- **report compiled with working directory**: `r getwd()`
- **germplasms:** `r paste(meta$germplasms, collapse=", ")`
- **timepoints:** `r paste(meta$timepoints, collapse=", ")`
- **treatments:** `r paste(meta$treatments, collapse=", ")`
- **number of replicates:** `r paste(meta$nrepeats, collapse=", ")`
- **number of used plates:** `r length(unique(manifest$plate))`
- **configuration:** `r paste(meta$name, collapse=", ")`

## Plate layout

The images will be interpreted like this, using `r readManifest('../..')[1,'plate']`
as an example

```{r image-layout, results="asis", echo=FALSE}
manifest <- subset(manifest, timepoint == meta$timepoints[1] & plate == manifest[1, 'plate'])
fi <- merge(manifest, meta$griddf)
fi$rn <- paste(fi$germplasm_region, fi$Sample_ID, fi$GERMPLASM, sep=",")
dd <- reshape2::dcast(fi, ROW~RANGE, value.var='rn')[, -1]
mark <- rep('', nrow(dd))
mark[ceiling(length(mark) / 2)] <- '<- Mark'
ds <- cbind(dd, mark)
rownames(ds) <- 1:nrow(ds)
colnames(ds) <- c(1:(ncol(ds) - 1), '')
print(xtable::xtable(ds), 'html',
      html.table.attributes="class='table table-hover'",
        include.rownames=FALSE)
```

and the plates should be sown like this using the same example

```{r plate-layout, results="asis", echo=FALSE}
mark <- rep('', nrow(dd))
mark[ceiling(length(mark) / 2)] <- 'Mark ->'
dd <- cbind(mark, dd[nrow(dd):1, ncol(dd):1])
rownames(dd) <- 1:nrow(dd)
colnames(dd) <- c('', 1:(ncol(dd) - 1))
print(xtable::xtable(dd), 'html',
      html.table.attributes="class='table table-hover'",
      include.rownames=FALSE)
```

## Plate sowing

Sowing order is indicated below and also given as
[sowing-order.csv](sowing-order.csv) and in long format with one row
per germplasm in [sowing-order-long.csv](sowing-order-long.csv), and
as short printable summary in
[sowing-summary.csv](sowing-summary.csv)

```{r sowing, results="asis", echo=FALSE}
mfs <- subset(readManifest('../..'), timepoint == meta$timepoints[1])

sowingdf <- plyr::ddply(mfs, 'plate', function(dd) {
  data.frame(plate=unique(gsub('\\..*$', '', dd$plate)),
             label=unique(dd$label),
             block=unique(dd$BLOCK),
             position=unique(dd$position),
             treatment=unique(dd$treatment),
             germplasms=paste(dd$GERMPLASM, collapse=','),
             regions=paste(dd$germplasm_region, collapse=','))
})
sowingdf <- sowingdf[with(sowingdf, order(block, position)),]
rownames(sowingdf) <- NULL
print(xtable::xtable(sowingdf), 'html',
      html.table.attributes="class='table table-hover'",
        include.rownames=FALSE)
write.csv(sowingdf, file='sowing-order.csv', row.names=FALSE)

cols <- c('plate', 'label', 'BLOCK', 'position', 'treatment', 'GERMPLASM',
          'germplasm_region')
sowingdfLong <- mfs[,cols]
sowingdfLong <- transform(sowingdfLong, plate=gsub('\\..*$', '', plate))
colnames(sowingdfLong) <-
  c('plate', 'label', 'block', 'position', 'treatment', 'germplasm', 'region')
sowingdfLong <- sowingdfLong[with(sowingdfLong, order(block, position)),]
rownames(sowingdfLong) <- NULL
write.csv(sowingdfLong, file='sowing-order-long.csv', row.names=FALSE)
```

A compact listing over which plates are to subjected to which treatment:

```{r sowing-summary, results="asis", echo=FALSE}
manifest <- readManifest('../..')
summary <- manifest[,c("plate","germplasm_region","treatment","GERMPLASM")]
fun <- function(x) paste(unique(x),collapse=",")

summary <- reshape2::melt(summary,
                          id.vars=c("plate", "germplasm_region", "treatment", "GERMPLASM"))
summary <- as.data.frame(reshape2::dcast(summary,
                                         GERMPLASM + germplasm_region ~ treatment,
                                         fun,
                                         value.var=as.character('plate')))
treatments <- unique(manifest$treatment)

for (i in treatments){
  plates <- plyr::ldply(summary[,i], function(x) {
    string <- gsub(".jpg", "", gsub("plate0*", "", x))
    s <- sort(unique(as.numeric((unlist(strsplit(string, ","))))))
    dif <- s[seq(length(s))][-1] - s[seq(length(s)-1)]
    new <- !c(0, dif == 1)
    cs <- cumsum(new)
    res <- vector(mode="list", max(cs))
    for(i in seq(res)){
      s.i <- s[which(cs == i)]
      if(length(s.i) > 2){
        res[[i]] <- paste(min(s.i), max(s.i), sep="-")
      } else {
        res[[i]] <- as.character(s.i)
      }
    }
    paste(unlist(unique(res)), collapse=",")
  })
  summary[,i] <- plates
}

summary <- summary[order(summary[,treatments[1]], summary[,'germplasm_region']),]
print(xtable::xtable(summary), 'html',
      html.table.attributes="class='table table-hover'",
      include.rownames=FALSE)
write.csv(summary, row.names=FALSE, file='sowing-summary.csv')
```
