---
title: "Introduction to *rosettR*"
author: Henning Redestig
email: henning.red@gmail.com
date: "22nd November 2015"
package: "`r packageVersion('rosettR')`"
abstract: The rosettR package implements a fully described, self contained, high-throughput and affordable phenotyping platform for measuring growth rate in Arabidopsis.
output:
  html_document
graphics: yes
vignette: >
  %\VignetteIndexEntry{Introduction to rosettR}
  %\VignettePackage{rosettR}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r pre-fix, echo=FALSE}
library(knitr)
opts_knit$set(upload.fun=image_uri)
```

# Introduction to *rosettR*

This package implements a high-throughput and affordable growth
phenotyping assay for plate grown Arabidopsis plants. Briefly, plants
are grown in petri dishes on solid medium and photographed at regular
intervals. Images are then analyzed using the functions implmented in
this package to calculate area and relative growth rate of the rosetts
which then can be used to compare the differences between genotypes
and the effect of applied treatments such as starvation, osmotic or
heat stress.

In this vignette, the creation of a new experiment, analysis of images
and generation of template statistical reports are described. The
examples are reproducible and with a working installation of `rosettR`
you can copy-paste code to your R terminal to try it out.

## Installation

This package is not yet available on CRAN so the easiest way to
install it is to use the [devtools](http://github.com/hadley/devtools)
package.

```{r eval=FALSE}
library(devtools)
install_github("hredestig/rosettR")
```

The package is tested on Windows 7 and Linux.

# Experiment design

A new experiment can be started by first loading the package and
creating a meta-data object that details the specifics of the
experiment. As a toy-example, let's study genotypes named foo, bar,
baz and qux subjected to two treatments: control condition and osmotic
stress. Each plate is in this design divided in four regions with 8
boxes each (aka /wells/) of which all are occupied by a single seed of
the same genotype. In our test experiment, we have only four genotypes
so all plates have the same layout. In addition to genotypes and
treatment, we define the time-points (days) at which pictures are
taken, the number of repetition blocks and short description of the
experiment to add to each report.

* **genotypes** the names of the studied genotype
* **treatments** the applied treatment conditions.
* **time-points** the considered time-points (days)
* **number of repetition blocks** the number of times the experiment
  is repeated to obtain biological replication
* **reference** the genotype with which all other genotypes should be
  compared 
* **description** a optional short description of what the experiment
  is meant for

*NB:* Avoid using special character such as greek letters. If
you must have spaces within an element, make sure to quote it as
`"genotype with space"`. 

We start an experiment by first creating a meta-data object (a simple
list with key-value pairs)

```{r setup}
library(rosettR)
meta <- metaTemplate(genotypes=c("foo", "bar", "baz", "qux"),
                     treatments=c("control", "osmotic_stress"),
                     timepoints=c(11, 14, 16, 18),
                     nblocks=3,
                     reference="foo",
                     description="a test experiment")
```

Next we 'create' the experiment which generates the metadata-file, the
directory structure in which to place the taken pictures and the
manifest file that defines the randomized block design.

```{r create-experiment}
pathToExperiment <- "rosettrTest"
newExperiment(pathToExperiment, meta)
```

The manifest file can be open in your favor spread-sheet editor or
just inspected directly by

```{r example-manifest}
head(readManifest(pathToExperiment))
```

To obtain an overview of the created experiment, to support preparing
the plates, compile the template 'layout' report.

```{r layout-report, results="hide"}
makeReport(pathToExperiment, "layout", quiet=TRUE)
```

Let plants grow, take images (in jpg format) exactly in the order they
are listed in the manifest and save them in the designated directories
for each monitored day. Make sure that your camera's internal watch is
correctly set as the time the photo was taken will be used to match
each image to the corresponding line in the manifest. You are further
adviced to lavel the plates (in the margin of the lid, visible to the
camera but not covering any plants) so that you can be sure the order
is correct. Further ensure that

* each plate is exactly in the centre of the image
* that it is straight, with the grid parallel to the edges of the
  image
* the background color and light is the same across all days
* that the zoom factor is the same across all days

Thanks to to auto-correction steps in the analysis, it is possible to
analyze image also with slight deviations from these rules but it
result is likely to be better if they are kept.

# Image analysis

Before detailing the process of analyzing a whole experiment, we give
a brief description of the analysis of a single image. This process is
defined in the function `analyzeImage` and consists of the
following main steps:

* image files are renamed from e.g. IMGxxxx.jpg to plate001.jpg to
  match the names in the manifest (order obtained from time the
  picture was taken as recorded in the exif tag of the jpeg file)
* re-size the image to the `hires` width - default to 1500 pixels
  width (smaller gives faster analysis but too small will introduce
  noise)
* test for eccentricity of the plate in which the plate is detected
  and its eccentricity is recorded and correcte for.
* adjustment for plate location (using an optimization algorithm)
* made binary by clustering
* adjustment for plate rotation (using a mixture model algorithm)
* extracting contiguous features
* sorting features to different boxes in the grid by checking
  their distances to box-centers and total occupancy of boxes.
* generating a quality control picture.

For testing purposes, we first unzip some example images to our
example and set the scale directly

```{r unzip-example}
meta <- readMeta(pathToExperiment)
meta$pixelsmm <- 7.54 # 7.54 pixels per millimeter
writeMeta(meta, pathToExperiment)
unzip(system.file("examples/rosettrTest.zip", package="rosettR"),
      exdir=".")
```

When all images have been taken you can generate an overview of them by

```{r overview}
makeReport(pathToExperiment, "overview", quiet=TRUE)
```

If all images look as expwcted, it is time to analyze them. First
however, you should calibrate the scale of the images to get the right
pixels to millimeter mapping. Do this by

```{r calibrate, eval=FALSE}
calibrateScale(pathToExperiment)
```

Which will record the conversion in the meta-data file for future
reference.

Next, we analyze the images using `processPlateExperiment` function.

```{r eval=FALSE} 
processPlateExperiment(pathToExperiment, checklocation=FALSE)
```
```{r echo=FALSE}
unzip(system.file("examples/rosettrTestResults.zip", package="rosettR"),
      exdir=".")
```

we turn off the plate location detection to speedup this test analysis (and
for demonstration purposed later in this vignette)

## Perform quality control

Once images have been analyzed, statistical analysis can be done by
compiling a suitable template report. This is done by running
`makeReport` that creates a report for the experiment
in the designated output directory (`[path to
    experiment]/Output/reports/`). The first report to compile is
typically a quality control report.

```{r qc-check, results="hide"}
makeReport(pathToExperiment, "quality-check", quiet=TRUE)
```

The final report can then be found in the `Output` folder of the
experiment which shows QC images for all plates as well as graphs to
indicate the presence of outliers.

In the QC images, the grid is indicated by black boxes and features
are shown with randomly chosen colors. Boxes in which features were
found that could not be sorted to a distinct box are indicated by a
red border. In a successful scenario, all boxes are black and each box
contains a plant of color different from the plants in all neighboring
boxes. However, when overlapping plants are identified, the
corresponding boxes are cutout and all the area estimate of the plant
in that box becomes the sum of all features in the box. This process
will often yield poor area estimates and plates with many ambiguous
boxes may be considered for removal using the `removeBoxes` function,
or reprocessed with tuned image analysis (arguments for
`analyzeImage`) parameters.

## Re-processing failed images

If the analysis has gone wrong due to e.g the plate location not
correctly identified, it is possible to reprocess the picture with
improved parameters to image threshold, angle correction, plate
displacement etc.

As can be seen when studying the QC images of our example experiment,
one image failed since it was excessively out of the centre of the
image.

```{r excentric}
pda <- readPhenodata(pathToExperiment)
qcpic <- subset(pda, timepoint == 11 & plate == "plate001.jpg")$qc[1]
library(EBImage)
qcPath <- file.path(pathToExperiment, qcpic)
qcImage <- readImage(qcPath)
display(qcImage, method="raster")
```

To resolve this problem, we reprocess that single image but this time
with the location correction turned on.

```{r fix, results="hide"}
mf <- readManifest(pathToExperiment)
mf <- subset(mf, timepoint == 11 & plate  == "plate001.jpg")
## reprocess the plate where using the plate location correction
pda <- reprocessPlateImages(pathToExperiment, mf, checklocation=TRUE)
```

Inspect the result

```{r inspect-fix}
pda <- readPhenodata(pathToExperiment)
qcpic <- subset(pda, timepoint == 11 & plate == "plate001.jpg")$qc[1]
display(readImage(file.path(pathToExperiment, qcpic)), method="raster")
```

See the documenation for `analyzeImage` for other parameters that can
be adjusted during image analysis.

We ca then re-compile the quality control report

```{r qc-again, results="hide"}
makeReport(pathToExperiment, "quality-check", quiet=TRUE)
```

# Interpretation

To obtain the final results from the experiment, compile the report to
compare areas by plots and by ANOVA.

```{r compare, results="hide"}
makeReport(pathToExperiment, "compare-areas", quiet=TRUE)
```

You can also simply run

```{r manual-analysis, eval=FALSE}
readPhenodata(pathToExperiment)
```

to get a dataframe with the results with which you can perform your
own tailored statistical analysis.

# Additional topics

## Failed plates

In the event of contamination, e.g. fungal growth on the plate you may
want to prematurely discard plates. In that case, it is important to
take place-holder images to get the order of images right. The
recommendation is to replace the bad plate with an empty one and keep
taking images of that empty plate. Such plates are detected by
counting the fraction between 'dark' pixels and light pixels. If this
fraction falls below 0.0001, the plate is considered empty and the
analysis will be skipped. The QC picture will be replace by a place
holder indicated that the plate was empty. See the experiment results
above for an example.

## Parallel computing

Image analysis can be done in parallel to speed up computation
considerably. The `rosettR` package uses
[plyr](https://cran.r-project.org/web/packages/plyr/index.html) which
support using multiple CPUs on a single host via the
[doParallel](https://cran.r-project.org/web/packages/doParallel/index.html)
package. Simple register a parallel backend by for using e.g. 4 CPUs
by

```{r eval=FALSE}
library(doParallel)
registerDoParallel(cores=4)
```

and then processing images adding the argument `.parallel=TRUE`. The
memory consumption is quite low so you can safely use as many CPUs as
you have minus 1 (to not make your computer unresponsive).

